---
title: "Creating BAF plots"
author: "Brian J. Knaus"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BAF plots}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


Use of B allele plots is intended to find aberrations in sequence depth and allele frequency.


## Preliminaries



## Workflow


In `vignette('intro_to_read_matrix')` we discussed the use of `read_matrix`.
Our first task is to read in the data.


```{r}
library(covR)
ex_file <- system.file("extdata", "sc10_4k.mpileup.gz", package = "covR")
```


We can now take our first pass through the file to collect information.

```{r}
stats <- file_stats(ex_file, verbose=0)
stats
x1 <- read_matrix(ex_file, nrows=stats['Rows'], cols=c(1:3, 5, 6), verbose=0)
head(x1)
```



This is mpileup format data.
The first three columns identify the chromosome, position and reference allele.
Samples occur in subsequent columns where each samples consists of three columns (here I've included the optional depth column).
Here, I've only read in the first sample and omitted the depth column.


We need to convert this data from the mpileup format to a table of counts.
This is accomplished with the function `baf_stats`.


```{r}
x2 <- baf_stats(x1)
head(x2)
```


We now have a table of counts for each nucleotide (forward and reverse in capitals and lowercase) as well as an asterix for deletions.
We can plot this information to visualize it with `baf_plot`.


```{r, fig.align='center', fig.height=4, fig.width=6}
baf_plot(x2)
```

## Future directions



## Resources

There are quite a diversity of softwares available for the analysis of copy number variation.
Some methods may be specific to a particular technology (e.g., microarrays, Illumina short reads, etc.).
Some methods may be specific to a particular organism, such as human.
When shopping for a software for your project, these will be an important consideration.
Here's a short list of resources I find interesting.


* [CopyNumberVariation](http://www.bioconductor.org/packages/release/BiocViews.html#___CopyNumberVariation) BiocView at Bioconductor

* [Coverage](http://www.bioconductor.org/packages/release/BiocViews.html#___Coverage) analyses at Bioconductor
* [genoCN](http://www.bioconductor.org/packages/release/bioc/html/genoCN.html) at Bioconductor
* [RDXplorer](http://rdxplorer.sourceforge.net/) - event-wise testing (EWT)
* [cn.mops](http://www.bioconductor.org/packages/release/bioc/html/cn.mops.html)
* [readDepth](https://github.com/chrisamiller/readdepth)
* [VarScan 2](http://varscan.sourceforge.net/)
* [SegSeq](http://www.broadinstitute.org/cgi-bin/cancer/publications/pub_paper.cgi?mode=view&paper_id=182)


[Olshen](http://biostatistics.oxfordjournals.org/content/5/4/557.short?rss=1&ssource=mfc) et al.'s description of circular binary segmentation (originally implemented on microarrays).


